---
title: "Terrestrial surveys"
format: html
---

# Terrestrial surveys

```{r}
library(tidyverse)
library(patchwork)
```


```{r}
raw_data <- read_csv("2023/data/final_herbivore-data/ng18_herbivore_survey_data.csv")
raw_data <- raw_data |>
  mutate(
    id = if_else(id == 4.1 & transect == "mopane", 4.2, id),
    count = if_else(count == 0, 1, count)
  )

functional_groups <- read_csv("shared/reference/mammal_functional_groups_terrestrial.csv")

raw_data <- raw_data |> left_join(functional_groups, by = "species")

# Get order and family names using `taxize`
higher_taxonomy <- taxize::tax_name(
  sci = c(raw_data$scientific_name |> unique()),
  get = c("order", "family"))

higher_taxonomy <- higher_taxonomy |>
  mutate(
    order = case_when(
      query == "Lupulella mesomelas" ~ "Carnivora",
      query == "Connochaetus taurinus" ~ "Artiodactyla",
      query == "Herpestes sanguineus" ~ "Carnivora",
      TRUE ~ order
    )
  )

raw_data <- raw_data |>
  left_join(higher_taxonomy, by = c("scientific_name" = "query"))
```



Total species richness and abundance per site

```{r}
total_summary <- raw_data |>
  group_by(transect) |>
  summarise(
    total_sr = unique(species) |> length(),
    total_a = sum(count)
  ) |>
  ungroup()

totals <- raw_data |>
  summarise(
    transect = "Both",
    total_sr = unique(species) |> length(),
    total_a = sum(count)
  )

total_summary <- total_summary |>
  bind_rows(totals) |>
  mutate(transect = factor(transect, levels = c("floodplain", "mopane", "Both")))
```


Plot

```{r}
sr_total_plot <-
  total_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = total_sr, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 25)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Species Richness",
    x = "Survey",
    title = "a) Total"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
    title_size = 16
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )

a_total_plot <-
  total_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = total_a, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 800)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Abundance",
    x = "Survey"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14
  ) +
  theme(legend.position = "none", plot.margin = margin(10, 10, 10, 10))

sr_total_plot + a_total_plot
```



Mean species richness and abundance per site (averaged across days surveyed)

```{r}
mean_summary <- raw_data |>
  group_by(transect, Sample.Label) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup() |>
  group_by(transect) |>
  summarise(
    mean_species_richness = mean(species_richness),
    sr_std = sd(species_richness) / sqrt(length(species_richness)),
    mean_abundance = mean(abundance),
    a_std = sd(abundance) / sqrt(length(abundance)),
  ) |>
  ungroup()

means <- mean_summary |>
  summarise(
    transect = "Both",
    sr_std = sd(mean_species_richness) / sqrt(length(mean_species_richness)),
    a_std = sd(mean_abundance) / sqrt(length(mean_abundance)),
    mean_species_richness = mean(mean_species_richness),
    mean_abundance = mean(mean_abundance),
  )

mean_summary <- mean_summary |>
  bind_rows(means) |>
  mutate(transect = factor(transect, levels = c("floodplain", "mopane", "Both")))
```


Plot

```{r}
sr_mean_plot <-
  mean_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = mean_species_richness, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  geom_errorbar(
    aes(
      ymin = mean_species_richness - sr_std,
      ymax = mean_species_richness + sr_std,
      x = transect),
    width = 0.2,
    linewidth = 0.5) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 15)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Species Richness",
    title = "b) Mean",
    x = "Survey"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
    title_size = 16
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )

a_mean_plot <-
  mean_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = mean_abundance, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  geom_errorbar(
    aes(
      ymin = mean_abundance - a_std,
      ymax = mean_abundance + a_std,
      x = transect),
    width = 0.2,
    linewidth = 0.5) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 200)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Abundance",
    x = "Survey"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14
  ) +
  theme(legend.position = "none", plot.margin = margin(10, 10, 10, 10))

sr_mean_plot + a_mean_plot

sr_total_plot + a_total_plot + sr_mean_plot + a_mean_plot
ggsave("2023/figures/species_richness_abundance_terrestrial.png", width = 8, height = 8)
```


Alternative point and range plots instead of bar plots

```{r}
herbivore_summary <- raw_data |>
  group_by(transect, Sample.Label) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup()

herbivore_summary_all <- raw_data |>
  filter(Sample.Label != 4) |>
  group_by(Sample.Label) |>
  summarise(
    transect = "Both",
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup()

herbivore_summary <- herbivore_summary |>
  bind_rows(herbivore_summary_all) |>
  mutate(transect = factor(transect, levels = c("floodplain", "mopane", "Both")))

sr_mean_plot <-
  herbivore_summary |>
  ggplot() +
  stat_summary(
    aes(x = transect, y = species_richness, colour = transect),
    fun = mean,
    geom = "pointrange",
    fun.min = min,
    fun.max = max,
    size = 0.5,
    linewidth = 1
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_colour_manual(values = c("#66C2A5", "#FC8D62", "#4a4a4a")) +
  labs(
    y = "Species Richness",
    x = "Survey",
    title = "b) Mean"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_mean_plot

a_mean_plot <-
  herbivore_summary |>
  ggplot() +
  stat_summary(
    aes(x = transect, y = abundance, colour = transect),
    fun = mean,
    geom = "pointrange",
    fun.min = min,
    fun.max = max,
    size = 0.5,
    linewidth = 1
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(50, 300)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_colour_manual(values = c("#66C2A5", "#FC8D62", "#4a4a4a")) +
  labs(
    y = "Abundance",
    x = "Survey"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
    title_size = 16
  ) +
  theme(legend.position = "none", plot.margin = margin(10, 10, 10, 10))
a_mean_plot

sr_mean_plot + a_mean_plot

sr_total_plot + a_total_plot + sr_mean_plot + a_mean_plot
ggsave("2023/figures/species_richness_abundance_terrestrial_v2.png", width = 8, height = 8)
```


Maximum daily species richness and abundance per transect and maximum cumulative species richness across both transects

```{r}
max_summary <- raw_data |>
  group_by(transect, Sample.Label) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup() |>
  group_by(transect) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance)
  ) |>
  ungroup()

max_overall <- raw_data |>
  group_by(Sample.Label) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup() |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance),
    transect = "Both"
  ) |>
  ungroup()

max_summary <- max_summary |>
  bind_rows(max_overall) |>
  mutate(transect = factor(transect, levels = c("floodplain", "mopane", "Both")))

sr_max_plot <-
  max_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = max_species_richness, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 18)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Maximum Species Richness",
    title = "a)",
    x = "Survey",
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_max_plot

a_max_plot <-
  max_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = max_abundance, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 300)
  ) +
  scale_x_discrete(
    expand = c(0.25, 0.25),
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Maximum Abundance",
    x = "Survey",
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_max_plot

sr_max_plot + a_max_plot
```


Box plots of daily species richness per transect and daily cumulative richness across all transects

```{r}
summary <- raw_data |>
  group_by(transect, Sample.Label) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup()

overall <- raw_data |>
  group_by(Sample.Label) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count),
    transect = "Both"
  ) |>
  ungroup() |>
  filter(Sample.Label != 4)

summary <- summary |>
  bind_rows(overall) |>
  mutate(transect = factor(transect, levels = c("floodplain", "mopane", "Both")))

sr_boxplot <-
  summary |>
  ggplot() +
  geom_boxplot(
    aes(y = species_richness, x = transect, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_x_discrete(
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 18)
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Species Richness",
    title = "b)",
    x = "Survey",
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_boxplot
```


Box plots of daily abundance for each site then daily abundance summing all sites

```{r}
a_boxplot <-
  summary |>
  ggplot() +
  geom_boxplot(
    aes(y = abundance, x = transect, fill = transect),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_x_discrete(
    labels = c("Floodplain", "Mopane", "Both")
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 300)
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#ededed")) +
  labs(
    y = "Abundance",
    x = "Survey",
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_boxplot

sr_boxplot + a_boxplot

sr_max_plot + a_max_plot + sr_boxplot + a_boxplot
ggsave("2023/figures/max_species_richness_abundance_and_boxplots_herbivores.png", width = 9, height = 9)
```


Calculate summaries of mammal orders

```{r}
orders_summary <- raw_data |>
  group_by(transect, order) |>
  summarise(
    total_sr = unique(species) |> length(),
    total_a = sum(count)
  ) |>
  ungroup()

orders_totals <- raw_data |>
  group_by(order) |>
  summarise(
    transect = "Both",
    total_sr = unique(species) |> length(),
    total_a = sum(count)
  ) |>
  ungroup()

orders_summary <- orders_summary |>
  complete(transect, order, fill = list(total_sr = 0, total_a = 0)) |>
  bind_rows(orders_totals) |>
  mutate(
    transect = factor(transect, levels = c("floodplain", "mopane", "Both")),
    order = factor(order, levels = c("Artiodactyla", "Perissodactyla", "Proboscidea", "Carnivora", "Primates", "Rodentia", "Struthioniformes"))
  )
```


Plot

```{r}
orders_sr_plot <-
  orders_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = total_sr, fill = order),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 12)
  ) +
  scale_x_discrete(labels = c("Floodplain", "Mopane", "Both")) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Species Richness",
    x = "Survey",
    fill = "Order"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 20, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "bottom"
  )

orders_a_plot <-
  orders_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = total_a, fill = order),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 450)
  ) +
  scale_x_discrete(labels = c("Floodplain", "Mopane", "Both")) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Abundance",
    x = "Survey",
    fill = "Order"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 10, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "bottom"
  )

design <- "
12
33
"

orders_sr_plot +
  orders_a_plot +
  guide_area() +
  patchwork::plot_layout(design = design, guides = "collect", heights = c(7, 1))
ggsave("2023/figures/orders_vertebrates_terrestrial.png", width = 12, height = 5.5)
```


Maximum daily richness and abundance per transect and maximum cumulative species richness across both transects

```{r}
orders_summary <- raw_data |>
  group_by(transect, Sample.Label, order) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup() |>
  group_by(transect, order) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance)
  ) |>
  ungroup()

orders_overall <- raw_data |>
  filter(Sample.Label != 4) |>
  group_by(Sample.Label, order) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup() |>
  group_by(order) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance),
    transect = "Both"
  ) |>
  ungroup()

orders_summary <- orders_summary |>
  bind_rows(orders_overall) |>
  complete(transect, order, fill = list(max_species_richness = 0, max_abundance = 0)) |>
  mutate(
    transect = factor(transect, levels = c("floodplain", "mopane", "Both")),
    order = factor(order, levels = c("Artiodactyla", "Perissodactyla", "Proboscidea", "Carnivora", "Primates", "Rodentia", "Struthioniformes"))
  )

orders_sr_plot <-
  orders_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = max_species_richness, fill = order),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_x_discrete(labels = c("Floodplain", "Mopane", "Both")) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 9)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Maximum Species Richness",
    x = "Survey",
    fill = "Order"
  ) +
  guides(fill = guide_legend(byrow = TRUE)) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 20, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top"
  )
orders_sr_plot

orders_a_plot <-
  orders_summary |>
  ggplot() +
  geom_col(
    aes(x = transect, y = max_abundance, fill = order),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_x_discrete(labels = c("Floodplain", "Mopane", "Both")) +
  scale_y_continuous(
    expand = c(0, 0),
    #limits = c(0, 500)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Maximum Abundance",
    x = "Survey",
    fill = "Order"
  ) +
  guides(fill = guide_legend(byrow = TRUE)) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 10, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top"
  )
orders_a_plot

design <- "
12
33
"

orders_sr_plot +
  orders_a_plot +
  guide_area() +
  patchwork::plot_layout(design = design, guides = "collect", heights = c(7, 1))
ggsave("2023/figures/max_richness_abundance_orders_herbivores.png", width = 12, height = 5.5)
```


Box plots of daily cumulative richness and abundance across all sites

```{r}
orders_summary <- raw_data |>
  filter(Sample.Label != 4) |>
  group_by(Sample.Label, order) |>
  summarise(
    species_richness = unique(species) |> length(),
    abundance = sum(count)
  ) |>
  ungroup() |>
  complete(Sample.Label, order, fill = list(species_richness = 0, abundance = 0))

sr_orders_boxplot <-
  orders_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = species_richness, x = order, fill = order),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 9)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Species Richness",
    x = "Order"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_orders_boxplot

a_orders_boxplot <-
  orders_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = abundance, x = order, fill = order),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 130)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Abundance",
    x = "Order"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_orders_boxplot

sr_orders_boxplot + a_orders_boxplot

ggsave("2023/figures/richness_abundance_boxplots_orders_herbivores.png", width = 17, height = 5.5)
```


Abundances by species

```{r}
species_abundance_summary <- raw_data |>
  group_by(common_name) |>
  summarise(species_abundance = sum(count)) |>
  ungroup()

species_abundance_summary |>
  ggplot() +
  geom_col(
    aes(x = common_name, y = species_abundance),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5,
    fill = "#ededed"
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 350)
  ) +
  labs(
    y = "Species Abundance",
    x = "Species"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 10,
    axis_title_size = 13
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

ggsave("2023/figures/terrestrial_surveys_species_abundance.png", width = 6, height = 5)
```


Abundances by species, grouped by transect

```{r}
species_abundance_transect_summary <- raw_data |>
  group_by(common_name, transect) |>
  summarise(species_abundance = sum(count)) |>
  ungroup() |>
  complete(common_name, transect, fill = list(species_abundance = 0))

species_abundance_transect_summary |>
  ggplot() +
  geom_col(
    aes(x = common_name, y = species_abundance, fill = transect),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 200)
  ) +
  scale_fill_brewer(
    palette = "Set2",
    labels = c("Floodplain", "Mopane")
  ) +
  labs(
    y = "Species Abundance",
    x = "Species",
    fill = "Survey"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 10,
    axis_title_size = 13,
    legend_text_size = 10,
    legend_title_size = 11
  ) +
  theme(
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top",
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

ggsave("2023/figures/terrestrial_surveys_species_abundance_transects.png", width = 6, height = 5.5)
```


Set up appendix table

```{r}
appendix_table <- raw_data |>
  group_by(transect, Sample.Label, common_name) |>
  summarise(
    abundance = sum(count)
  ) |>
  ungroup() |>
  group_by(transect, common_name) |>
  summarise(
    maximum_abundance = max(abundance)
  ) |>
  ungroup() |>
  group_by(common_name) |>
  summarise(
    transects_observed = paste0(transect, collapse = ", ") |> str_to_sentence(),
    maximum_abundance = max(maximum_abundance)
  ) |>
  ungroup() |>
  left_join(functional_groups) |> # Contains scientific names
  left_join(higher_taxonomy, by = c("scientific_name" = "query")) |>
  arrange(order, scientific_name) |>
  select(common_name, scientific_name, transects_observed, maximum_abundance, order)

#appendix_table |> write_csv("2023/data/terrestrial_appendix_table.csv")
```