---
title: "Riparian surveys"
format: html
---

# Riparian surveys

```{r}
library(tidyverse)
library(readxl)
library(patchwork)
```


Read in data

```{r}
riparian_survey_data <- read_xlsx("2023/data/waterbirds/bird_watching_data.xlsx")
```


## Birds

Filter to birds

```{r}
bird_data <- riparian_survey_data |>
  dplyr::filter(Type == "Bird", Species != "Ibis") |> # Remove "Ibis"
  mutate(
    Species = if_else(Species == "Harrier", "Pallid Harrier", Species), # Replace harrier with pallid harrier as Richard requested
    survey_day = str_extract(ID, "\\d")
  )

bird_functional_groups <- read_csv("shared/reference/bird_functional_groups.csv")

bird_data <- bird_data |> left_join(bird_functional_groups, by = "Species")
```


Total species richness and abundance per site across all days

```{r}
bird_data_total_summary <- bird_data |>
  group_by(Site) |>
  summarise(
    total_sr = unique(Species) |> length(),
    total_a = sum(Count)
  ) |>
  ungroup()

bird_data_totals <- bird_data |>
  summarise(
    Site = "All",
    total_sr = unique(Species) |> length(),
    total_a = sum(Count)
  )

bird_data_total_summary <- bird_data_total_summary |>
  bind_rows(bird_data_totals) |>
  mutate(Site = factor(Site, levels = c("A", "B", "C", "All")))
```


Plot

```{r}
sr_total_plot <-
  bird_data_total_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = total_sr, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 50),
    breaks = seq(0, 50, by = 25)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(
    y = "Species Richness",
    title = "a) Total"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
    title_size = 16
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )

a_total_plot <-
  bird_data_total_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = total_a, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 300)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(y = "Abundance") +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14
  ) +
  theme(legend.position = "none", plot.margin = margin(10, 10, 10, 10))

sr_total_plot + a_total_plot
```


Mean species richness and abundance per site (averaged across days surveyed) and mean species richness and abundance averaged across sites

```{r}
bird_data_mean_summary <- bird_data |>
  group_by(Site, ID) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(Site) |>
  summarise(
    mean_species_richness = mean(species_richness),
    sr_std = sd(species_richness) / sqrt(length(species_richness)),
    mean_abundance = mean(abundance),
    a_std = sd(abundance) / sqrt(length(abundance)),
  ) |>
  ungroup()

bird_data_means <- bird_data_mean_summary |>
  summarise(
    Site = "All",
    sr_std = sd(mean_species_richness) / sqrt(length(mean_species_richness)),
    a_std = sd(mean_abundance) / sqrt(length(mean_abundance)),
    mean_species_richness = mean(mean_species_richness),
    mean_abundance = mean(mean_abundance),
  )

bird_data_mean_summary <- bird_data_mean_summary |>
  bind_rows(bird_data_means) |>
  mutate(Site = factor(Site, levels = c("A", "B", "C", "All")))
```


Plot

```{r}
sr_mean_plot <-
  bird_data_mean_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = mean_species_richness, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  geom_errorbar(
    aes(
      ymin = mean_species_richness - sr_std,
      ymax = mean_species_richness + sr_std,
      x = Site),
    width = 0.2,
    linewidth = 0.5) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 15)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(
    y = "Species Richness",
    title = "b) Mean"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )

a_mean_plot <-
  bird_data_mean_summary |>
  ggplot() +
  geom_col(aes(x = Site, y = mean_abundance, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  geom_errorbar(
    aes(
      ymin = mean_abundance - a_std,
      ymax = mean_abundance + a_std,
      x = Site),
    width = 0.2,
    linewidth = 0.5) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 35)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(y = "Abundance") +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
    title_size = 16
  ) +
  theme(legend.position = "none", plot.margin = margin(10, 10, 10, 10))

sr_mean_plot + a_mean_plot
```


Alternative to bar plots

```{r}
bird_data_summary <- bird_data |>
  group_by(Site, survey_day) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup()

bird_data_summary_all <- bird_data |>
  group_by(survey_day) |>
  summarise(
    Site = "All",
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup()

bird_data_summary <- bird_data_summary |>
  bind_rows(bird_data_summary_all) |>
  mutate(Site = factor(Site, levels = c("A", "B", "C", "All")))

sr_mean_plot <-
  bird_data_summary |>
  ggplot() +
  stat_summary(
    aes(x = Site, y = species_richness, colour = Site),
    fun = mean,
    geom = "pointrange",
    fun.min = min,
    fun.max = max,
    size = 0.5,
    linewidth = 1
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_colour_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#4a4a4a")) +
  labs(
    y = "Species Richness",
    title = "b) Mean"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_mean_plot

a_mean_plot <-
  bird_data_summary |>
  ggplot() +
  stat_summary(
    aes(x = Site, y = abundance, colour = Site),
    fun = mean,
    geom = "pointrange",
    fun.min = min,
    fun.max = max,
    size = 0.5,
    linewidth = 1
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 90)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_colour_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#4a4a4a")) +
  labs(y = "Abundance") +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
    title_size = 16
  ) +
  theme(legend.position = "none", plot.margin = margin(10, 10, 10, 10))
a_mean_plot

sr_mean_plot + a_mean_plot

sr_total_plot + a_total_plot + sr_mean_plot + a_mean_plot
ggsave("2023/figures/species_richness_abundance_birds_v2.png", width = 7, height = 7)
```


Maximum species richness across survey dates and maximum cumulative species richness across all sites across survey dates (?)

```{r}
bird_data_max_summary <- bird_data |>
  group_by(Site, survey_day) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(Site) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance)
  ) |>
  ungroup()

bird_data_max_overall <- bird_data |>
  group_by(survey_day) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance),
    Site = "All"
  ) |>
  ungroup()

bird_data_max_summary <- bird_data_max_summary |>
  bind_rows(bird_data_max_overall) |>
  mutate(Site = factor(Site, levels = c("A", "B", "C", "All")))

sr_max_plot <-
  bird_data_max_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = max_species_richness, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 30)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(
    y = "Maximum Species Richness",
    title = "a)"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_max_plot
```


Maximum abundance across survey dates and maximum cumulative abundance across all sites across survey dates

```{r}
a_max_plot <-
  bird_data_max_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = max_abundance, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 90)
  ) +
  scale_x_discrete(expand = c(0.25, 0.25)) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(
    y = "Maximum Abundance"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_max_plot

sr_max_plot + a_max_plot
```


Box plots of daily species richness per site and daily cumulative richness across all sites

```{r}
bird_data_summary <- bird_data |>
  group_by(Site, survey_day) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup()

bird_data_overall <- bird_data |>
  group_by(survey_day) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count),
    Site = "All"
  ) |>
  ungroup()

bird_data_summary <- bird_data_summary |>
  bind_rows(bird_data_overall) |>
  mutate(Site = factor(Site, levels = c("A", "B", "C", "All")))

sr_boxplot <-
  bird_data_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = species_richness, x = Site, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 30)
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(
    y = "Species Richness",
    title = "b)"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_boxplot
```


Box plots of daily abundance for each site then daily abundance summing all sites

```{r}
a_boxplot <-
  bird_data_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = abundance, x = Site, fill = Site),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 90)
  ) +
  scale_fill_manual(values = c("#66C2A5", "#FC8D62", "#8DA0CB", "#ededed")) +
  labs(
    y = "Abundance"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_boxplot

sr_boxplot + a_boxplot

sr_max_plot + a_max_plot + sr_boxplot + a_boxplot
ggsave("2023/figures/max_species_richness_abundance_and_boxplots_birds.png", width = 9, height = 9)
```


Calculate summaries of feeding functional groups

```{r}
bird_functional_groups_summary <- bird_data |>
  group_by(Site, survey_day, functional_group) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(Site, functional_group) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance)
  ) |>
  ungroup()

bird_functional_groups_totals <- bird_data |>
  group_by(survey_day, functional_group) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(functional_group) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance),
    Site = "All"
  ) |>
  ungroup()

bird_functional_groups_summary <- bird_functional_groups_summary |>
  bind_rows(bird_functional_groups_totals) |>
  mutate(Site = factor(Site, levels = c("A", "B", "C", "All")))
```


Plot

```{r}
functional_groups_sr_plot <-
  bird_functional_groups_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = max_species_richness, fill = functional_group),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 15)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Maximum Species Richness",
    fill = "Functional Group"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 20, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top"
  )

functional_groups_a_plot <-
  bird_functional_groups_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = max_abundance, fill = functional_group),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 50)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Maximum Abundance",
    fill = "Functional Group"
  ) +
  guides(fill = guide_legend(byrow = TRUE)) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 10, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top"
  )

design <- "
12
33
"

functional_groups_sr_plot +
  functional_groups_a_plot +
  guide_area() +
  patchwork::plot_layout(design = design, guides = "collect", heights = c(7, 1))
ggsave("2023/figures/max_richness_abundance_functional_groups_birds.png", width = 12, height = 5.5)
```


Box plots

```{r}
bird_functional_groups_summary <- bird_data |>
  group_by(survey_day, functional_group) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup()

sr_functional_groups_boxplot <-
  bird_functional_groups_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = species_richness, x = functional_group, fill = functional_group),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 13)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Species Richness",
    x = "Functional Group"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_functional_groups_boxplot

a_functional_groups_boxplot <-
  bird_functional_groups_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = abundance, x = functional_group, fill = functional_group),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 50)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Abundance",
    x = "Functional Group"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_functional_groups_boxplot

sr_functional_groups_boxplot + a_functional_groups_boxplot

ggsave("2023/figures/richness_abundance_boxplots_functional_groups_birds.png", width = 12, height = 5.5)
```


## Other vertebrates

Tidy data

```{r}
mammal_reptile_data <- riparian_survey_data |>
  filter(Type %in% c("Mammal", "Reptile"))

mammal_reptile_functional_groups <- read_csv("shared/reference/mammal_reptile_functional_groups_riparian.csv")

mammal_reptile_data <- mammal_reptile_data |>
  left_join(mammal_reptile_functional_groups, by = "Species") |>
  mutate(survey_day = str_extract(ID, "\\d"))
```


Summarise data into functional groups

```{r}
verts_functional_groups_summary <- mammal_reptile_data |>
  group_by(Site, survey_day, functional_group) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(Site, functional_group) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance)
  ) |>
  ungroup()

verts_functional_groups_overall <- mammal_reptile_data |>
  group_by(survey_day, functional_group) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(functional_group) |>
  summarise(
    max_species_richness = max(species_richness),
    max_abundance = max(abundance),
    Site = "All"
  ) |>
  ungroup()

verts_functional_groups_summary <- verts_functional_groups_summary |>
  complete(Site, functional_group, fill = list(max_species_richness = 0, max_abundance = 0)) |>
  bind_rows(verts_functional_groups_overall) |>
  mutate(
    Site = factor(Site, levels = c("A", "B", "C", "All")),
    functional_group = factor(functional_group, levels = c("Herbivore", "Carnivore", "Omnivore"))
  )
```


```{r}
vert_functional_groups_sr_plot <-
  verts_functional_groups_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = max_species_richness, fill = functional_group),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    #limits = c(0, 7)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Maximum Species Richness",
    fill = "Functional Group"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 20, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top"
  )

vert_functional_groups_a_plot <-
  verts_functional_groups_summary |>
  ggplot() +
  geom_col(
    aes(x = Site, y = max_abundance, fill = functional_group),
    position = position_dodge2(padding = 0),
    colour = "#141414",
    width = 0.7,
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    #limits = c(0, 500)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Maximum Abundance",
    fill = "Functional Group"
  ) +
  guides(fill = guide_legend(byrow = TRUE)) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 18,
    axis_title_size = 20,
    legend_text_size = 18,
    legend_title_size = 19
  ) +
  theme(
    plot.margin = margin(10, 10, 10, 10),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "top"
  )

design <- "
12
33
"

vert_functional_groups_sr_plot +
  vert_functional_groups_a_plot +
  guide_area() +
  patchwork::plot_layout(design = design, guides = "collect", heights = c(7, 1))
ggsave("2023/figures/max_richness_abundance_functional_groups_mammals_reptiles.png", width = 12, height = 5.5)
```


Box plots

```{r}
verts_functional_groups_summary <- mammal_reptile_data |>
  group_by(survey_day, functional_group) |>
  summarise(
    species_richness = unique(Species) |> length(),
    abundance = sum(Count)
  ) |>
  ungroup() |>
  complete(survey_day, functional_group, fill = list(species_richness = 0, abundance = 0))

sr_functional_groups_boxplot <-
  verts_functional_groups_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = species_richness, x = functional_group, fill = functional_group),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 7)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Species Richness",
    x = "Functional Group"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
sr_functional_groups_boxplot

a_functional_groups_boxplot <-
  verts_functional_groups_summary |>
  ggplot() +
  geom_boxplot(
    aes(y = abundance, x = functional_group, fill = functional_group),
    width = 0.6,
    colour = "#141414",
    linewidth = 0.5
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 140)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    y = "Abundance",
    x = "Functional Group"
  ) +
  pilot::theme_pilot(
    axes = "lb",
    grid = "h",
    axis_text_size = 12,
    axis_title_size = 14,
  ) +
  theme(
    legend.position = "none",
    plot.title.position = "plot",
    plot.title = element_text(margin = margin(0, 0, 20, 0)),
    plot.margin = margin(10, 20, 10, 10)
  )
a_functional_groups_boxplot

sr_functional_groups_boxplot + a_functional_groups_boxplot

ggsave("2023/figures/richness_abundance_boxplots_functional_groups_mammals_reptiles.png", width = 12, height = 5.5)
```


Tidy bird data for appendix table

```{r}
bird_table <- bird_data |>
  group_by(Site, ID, Species) |>
  summarise(
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(Site, Species) |>
  summarise(
    maximum_abundance = max(abundance)
  ) |>
  ungroup() |>
  group_by(Species) |>
  summarise(
    sites_observed = paste0(Site, collapse = ", "),
    maximum_abundance = max(maximum_abundance)
  ) |>
  ungroup() |>
  left_join(bird_functional_groups) |>
  mutate(
    Species = str_to_sentence(Species),
    scientific_name = auk::ebird_species(Species, type = "scientific")
  ) |>
  arrange(functional_group)

bird_table |> filter(is.na(scientific_name)) |> pull(Species)

# Can fix NAs here but I did it manually in Excel, then sorted data alphabetically in Excel as well.

#bird_table |> write_csv("2023/data/birds_appendix_table.csv")
```


Tidy mammal and reptile data for appendix table

```{r}
mammal_reptile_table <- mammal_reptile_data |>
  group_by(Site, ID, Species) |>
  summarise(
    abundance = sum(Count)
  ) |>
  ungroup() |>
  group_by(Site, Species) |>
  summarise(
    maximum_abundance = max(abundance)
  ) |>
  ungroup() |>
  group_by(Species) |>
  summarise(
    sites_observed = paste0(Site, collapse = ", "),
    maximum_abundance = max(maximum_abundance)
  ) |>
  ungroup() |>
  left_join(mammal_reptile_functional_groups) |>
  mutate(Species = str_to_sentence(Species)) |>
  arrange(functional_group)

#mammal_reptile_table |> write_csv("2023/data/riparian_mammals_reptiles_appendix_table.csv")
```
